### 2016-03-29 chapters/background.tex
Added/solved reviews part 2/2

In the scope of supporting the configuration of a given product line some specific tools and techniques are used. This section gives an overview of what is used in this work to archive a simplification of the workflow from a "product line"-like source code base to a final product.

\subsection{Feature Models} \label{ch:fm}
Feature models are multi-purpose structure for product lines. 
\begin{figure}
	\includegraphics{img/img-fm.pdf}
	\caption{A simple example of a feature model}
	\label{img-fm}
\end{figure}
One of their benefits is the visualization as a feature diagram, providing an overview of the containing features, and their hierarchy and dependencies. In addition, it classifies the features and their dependencies by Ordinality (Optional/mandatory), logical operator (Disjunction/exclusive disjunction) and whether it is abstract or concrete. Furthermore, it contains all information needed to provide a configuration of Features and its validation.

Although feature models can be represented as a propositional expression in CNF, it is typically represented in form of a feature diagram. The feature diagram is structured as a tree. In that manner feature models map the hierarchy of features. The possible relationships of features with a common parent feature are \textit{or}, \textit{alternative} and \textit{and}. In the case of an \textit{or}-group, at least one of the child features has to be selected if the parent feature is selected. Within an \textit{alternative}-group exactly one of the child feature has to be selected if the parent feature is selected. When grouped with an \textit{and}-relation, any number of child features can be selected, if the parent feature is selected. In addition, child features in an \textit{and}-group can be marked as either mandatory or optional.  As features' relations may be of higher complexity than just parent-child relations additional constraints can be noted within a feature model. Constraints can contain any propositional expression.

Figure~\ref{img-fm} shows an example of a simple feature model. The node \textit{GraphLibrary} is the root feature. It's childs are the node \textit{Edges}, that has to be selected due to the mandatory marker, and \textit{Algorithms} that is marked optional. The implemented edge types are \textit{Directed} and \textit{Undirected} from which exactly one has to be chosen as they are mutually exclusive. From the algorithms at least one has to be chosen whenever \textit{Algorithms} is selected. The constraint beneath Figure~\ref{img-fm} implies that if the feature \textit{Cycle} is selected, the edge type needs to be directed.

\subsection{Product configuration}
The variability of a product line is represented by it's feature model, which itself is a set of features with specific interrelations. The process of configuration of a product line describes the steps to derive a product from the product line. To archive this, a user has to select a subset of all the possible features within the product line to meet his requirements. However, not all subsets of features result in a valid product. The interrelations of the features restrict the possible combinations of features resulting in a \textit{valid} configuration. If only one of the requirements from the interrelations between the features is validated, no product can be created and the configuration as such is considered \textit{invalid}. For automatic detection of validity, check the next section.

A non-final configuration, that is a configuration that still has some feature choices left to be made, is called a \textit{partial configuration}. A \textit{valid} partial configuration can be useful, as it narrows down the number of possible resulting variations and choices left to make.

During configuration a user selects or deselects features to his needs. This process requires a lot of domain knowledge on the one hand and detailed information about the implementation of each single feature on the other. With growing numbers of features and thus growing numbers of possible products, configuration requires more effort. However, the even greater problem arises from the possible interactions and interrelations of features, which one has to keep track of during configuration.

For all the interrelations of features which are contained within the feature model, an automated observation of the resulting constraints supports configuration. It does so by checking for satisfiability after each change in the configuration.

@@

The formalism of propositional expressions (see Section~\ref{ch:fm} allows feature models and configurations to be checked for validity. The propositional expression of a (partial) configuration is created on the basis of the propositional expression of the underlying feature model. Each selected feature is appended with a logical \textit{AND} and each specifically unselected feature is also appended with a logical \textit{AND} but gets negated. The resulting expression is then evaluated by a SAT-solver to check for satisfiability \cite{sat-solve}.

Even during configuration this process can be applied to check for invalid partial configurations after each decision. If a partial configuration reduces the valid choices in a single group to a single one, it can be made automatically. This is called \textit{propagation}. The automatic selection or deselection of a features due to propagation of a made selection always results in a valid configuration. Nevertheless the propagated selections can be confusing for the user.

>>>>

In the scope of supporting the configuration of a given product line some specific tools and techniques are used. This section gives an overview about the used techniques.

\subsection{Feature Models} \label{ch:fm}
\begin{figure}
	\includegraphics{img/img-fm.pdf}
	\caption{A simple example of a feature model}
	\label{img-fm}
\end{figure}

Feature models are a multi-purpose structure for product lines. One of their benefits is the visualization as a feature diagram, providing an overview of the containing features, and their hierarchy and dependencies. In addition, it classifies the features and their dependencies by ordinality, logical operator and whether it is abstract or concrete. Furthermore, it contains all information needed to provide a configuration of features and its validation. A configuration, or chosen set of features, is valid if there are no contradictions in the models dependencies or logical operators.

There are two main representations of feature models. The first approach is a propositional expression, which is a type of syntactic formula often used in propositional logic. However feature models are typically visualized in form of a feature diagram, which is structured as a tree \cite{fmgp}. In that manner feature models map the hierarchy of features. The possible relationships of features with a common parent feature are \textit{or}, \textit{alternative} and \textit{and}. All child features beneath a parent feature have the same relationship and therefore form a \textit{group}. In the case of an \textit{or}-group, at least one of the child features has to be selected if the parent feature is selected. Within an \textit{alternative}-group exactly one of the child feature has to be selected if the parent feature is selected. When features are grouped with an \textit{and}-relation, they are marked as either mandatory or optional. If the parent feature of an \textit{and}-group is selected, any number of optional marked features can be selected. Features marked as mandatory have to be selected. As features' relations may be of higher complexity than just parent-child relations additional constraints can be noted within a feature model. Constraints can contain any propositional expression.

In Figure~\ref{img-fm} we demonstrate an example of a simple feature model. The node \textit{GraphLibrary} is the root feature. It's childs are the node \textit{Edges}, that has to be selected due to the mandatory property, and \textit{Algorithms} that is marked optional. The implemented edge types are \textit{Directed} and \textit{Undirected} from which exactly one has to be chosen as they are mutually exclusive. From the algorithms at least one feature has to be chosen whenever \textit{Algorithms} is selected. The constraint at the bottom of Figure~\ref{img-fm} implies that if the feature \textit{Cycle} is selected, the edge type needs to be directed.

\subsection{Product Configuration}
The variability of a product line is represented by it's feature model, which itself is a set of features with specific interrelations. The configuration process of a product line describes the steps to derive a product from the product line. To archive this, a user has to select a subset of all the possible features within the product line to meet his requirements \cite{vrt}{\color{red} (nicht in Literaturliste)}. However, not all subsets of features result in a valid product. The interrelations of the features restrict the possible combinations of features resulting in a \textit{valid} configuration. {\color{red} Bin mir nicht sicher, was du damit sagen willst... umformulieren?} If only one of the requirements from the interrelations between the features is validated, no product can be created and the configuration as such is considered \textit{invalid}. For an automatic validity check, see the next section.

A non-final configuration, that is a configuration that still has some feature choices left to be made, is called a \textit{partial configuration}. A \textit{valid} partial configuration can be useful, as it narrows down the number of possible resulting variations and choices left to make. As the aim is to achieve a valid final configuration, there has to be a valid partial configuration after every choice made.

@@

The formalism of propositional expressions (see Section~\ref{ch:fm}) allows feature models and configurations to be checked for validity. Each selected feature is appended with a logical \textit{AND} and each specifically unselected feature is also appended with a logical \textit{AND} but gets negated. The resulting expression is then evaluated by a SAT-solver to check for satisfiability \cite{sat-solve}. If the expression is satisfiable the selected features make up a valid configuration.

Even during configuration this process can be applied to check for invalid partial configurations after each decision. If a partial configuration reduces the valid choices in a group to a single one, this choice can be made automatically. This is called \textit{propagation}. The automatic selection or deselection of a features due to propagation of a made selection always results in a valid configuration \cite{fmi}.

### 2016-03-29 chapters/intro.tex
Added/solved reviews part 2/2

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
Developing feature oriented, although bringing with it some overhead, facilitates the creation of software when it is planned on custom tailoring it for a multiple number of clients \cite{spl-pp}. Feature oriented means developing and maintaining single features that can be combined with others to create a whole product. All the individual features make up a product line, whereas a subset of those features create a variant of this product. Developing software product lines can result in a large amount of variants, when customizing the software to each customers needs. By developing with a feature-oriented approach the configuration of a single variant can be done by selecting the features a customer needs, automatically including its dependencies. The configuration is based on a feature model, that defines the available features and its relations to one another. A feature model documents the features of a product line and their relationships\cite{fospl}.

Feature Models are essential structures for the feature-oriented development and later configuration of software product lines (SPL). There are projects being developed in a feature-oriented manner, but don't have a feature model yet. Implementing it on top of a given Project can result in a complex task due to the amount of its features and the constraints between them. Nevertheless, the benefits of a correct feature model justify the effort to extract one out of a live project. This work aims at automatically generating it out of descriptive files and naming conventions, to simplify a big part of the feature model creation.

@@

FeatureIDE is an open source IDE for feature-oriented-software development. It provides all the functionality needed to programmatically generate and work with feature models. This includes the logic behind configurations, data structures and the visualization and processing of feature models. The underlying Eclipse enables us to implement the feature extraction and the questionnaire as a plugin.

This work demonstrates the feature model extraction and questionnaire creation based on a real life Project. The Project, an open source ERP system called \textit{Odoo}\footnote{https://www.odoo.com/} is predestined to be developed feature-oriented due to a huge amount of features and complex constraints. It is currently being developed and structured feature oriented, although it doesn't contain a feature model yet. Furthermore, \textit{Odoo} reached a critical amount of features where a salesman cannot consult a customer by going through all the features anymore, but has to come up with a more effective way.

Contributions:
\begin{itemize}
\item Automatic feature model generation based on descriptive files and naming conventions as a foundation for the thereon based questionnaire.
\item Eclipse plugin, that enables the creation of a questionnaire based on a configuration file and the associated feature model. The liberties of the configuration file specification allows very personalized questionnaires.
\item The questionnaire changes as choices are made based on the logic behind the configuration file and the dependencies of the feature model. This enables a responsive questionnaire that prevents unnecessary questions.
\item Exiting the questionnaire always results in a valid configured variant to always ensure a valid configuration.
\end{itemize}

>>>>

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
Developing feature oriented fashion facilitates the creation of software when it is planned on custom tailoring it for a multiple number of clients \cite{spl-pp}. Feature oriented means developing and maintaining single features that can be combined with others to create a whole product. All the individual features make up a product line, whereas a subset of those features create a variant of this product line. Developing software product lines can result in a large amount of variants, when customizing the software to each customers needs. By developing with a feature-oriented approach the configuration of a single variant can be done by selecting the features a customer needs. Often features have dependencies to one another, which can be added to a feature model to verify a configuration. A feature model documents the features of a product line and their relationships\cite{fospl}.

There are projects being developed in a feature-oriented manner, but don't have a feature model yet. Implementing it on top of a given project can result in a complex task due to the amount of its features and the constraints between them. Nevertheless, the benefits of a correct feature model justify the effort to extract one out of a live project. To simplify a big part of the feature model creation, this work aims at automatically generating it out of descriptive files and naming conventions.

@@

For the implementation of the feature extraction and the questionnaire, we used the eclipse plugin FeatureIDE. FeatureIDE is an open source IDE for feature-oriented-software development. It provides all the functionality needed to programmatically generate and work with feature models. This includes the logic behind configurations, data structures and the visualization and processing of feature models.

In the second part of this work, we demonstrate the feature-model extraction and questionnaire creation based on a real life project. The project, an open source ERP system called \textit{Odoo}\footnote{https://www.odoo.com/} is predestined to be developed feature-oriented due to a huge amount of features and complex constraints. It is currently being developed and structured feature oriented, although it doesn't contain a feature model yet. Furthermore, \textit{Odoo} reached a critical amount of features where a salesman cannot consult a customer by going through all the features anymore, but has to come up with a more effective way.

Contributions:
\begin{itemize}
\item Automatic feature-model generation based on descriptive files and naming conventions as a foundation for the thereon based questionnaire.
\item We introduce an Eclipse plugin, that enables the creation of a questionnaire based on a configuration file and the associated feature model. The liberties of the configuration file specification allows very personalized questionnaires.
\item The questionnaire reacts to the choices made by reacting differently on each questions outcome. This enables a responsive questionnaire that prevents unnecessary questions. In addition, every choice made creates a valid partial configuration, even if the questionnaire is exited early.
\item Illustration of the described workflow by applying it to a real world project.
\end{itemize}
